Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-02-15 doc/GMP2015/revision_1/s02_preliminaries.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

Other generic constraint system or (hyper)graph properties are the
following:
% (\note we refer to these as properties of the constraint
% system or as properties of the underlying (hyper)graph
% interchangeably).
A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate
rigidity matrix of indeterminates has independent rows (i.e, the
determinant of some square submatrix is not identically zero).
%It is {\em rigid} if the number of independent rows of the rigidity matrix is
%at least the number of columns less the number of rigid body motions,
%which is 3 for distance constraint systems.
It is \dfn{isostatic (minimally rigid, well-constrained)} if it is
both rigid and independent.
%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$,
%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
It is \dfn{flexible} if it is not rigid, \dfn{under-constrained} if it
is independent and not rigid, or \dfn{over-constrained} if it is not
independent.

@@

\subsection{Decomposition-Recombination (DR) Plans}
%
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{over-constrained} if it is not independent.
%
%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
%
%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
%
% \begin{definition}
%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
%
%
%
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
    following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced
        subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
        \item A leaf node is a single vertex.
A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$
        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
        and in fact, $C$ contains all the edges in the union of its
        children as well.
        We denote a complete DR-plan of $G$ as $\comdrp{G}$
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
    The maximum  fan-in is called the \dfn{size} of the DR-plan.
    We denote an optimal DR-plan of $G$ as $\optdrp{G}$.

@@

% \begin{definition}
%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$. By this, rule 3 is implicit for a complete DR-plan.
% \end{definition}

@@

%\emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.
% \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages. It is not useful to solve for two bodies simultaneously if there are no constraints between them.
% THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.

@@

\header{Realizing given qusecs.}
Note that this basic task
would require finding real solutions a large multivariate polynomial
system (of inequalities and equalities representing the constraints).
Without DR-planning, this requires double exponential time in the
number of variables (even if orientation type is specified). With
DR-planning, the complexity is dominated by the size of the largest
subsystem that is solved, or recombined from the solutions of its
child subsystems i.e.,the maximum fan-in occurring in a DR-plan. Thus
finding an optimal DR-plan is important for realization.

\header{Recursive block
decomposition of the rigidity matrix, stress space and external stresses.}
This can be obtained using the DR-plan we obtain a
with each block being the
rigidity matrix of an isostatic subgraph. The blocks intersect on
isostatic or trivial subgraphs. This recursive block decomposition of
the rigidity matrix $R$ correspondingly recursively decomposes
the stress vectors $s$ that balance a given external stress $t$ since
$sR = t$. Moreover, the external stress vector $t$ is
recursively ``distributed'' into external stresses for each
node of the DR-plan.
When the rigidity matrix and stress
vectors contain indeterminates, this  can be used to design
underlying graph and parameters $\delta$ to obtain an optimal
distribution of stresses on the geometric primitives. When the
realization of qusecs is given, this can be used to analyze the
distribution of a given external load.

@@

%%    \item \textbf{over-constrained} if it is not independent.

@@

%%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.

@@

%%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.

@@

%%    \item \textbf{under-constrained} if $G$ is independent and not rigid.

@@

%%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.

@@

%\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.

@@

%%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some

@@

%%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all

@@

%%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.

@@

%        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.

@@

%    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.

@@

%    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$. This makes rule 3 implicit. The complete DR-plan is unique.

@@

%%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$. By this, rule 3 is implicit for a complete DR-plan.

@@

%In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.

@@

%$C$ is the graph in an arbitrary node in $CompleteDRP(G)$. $C_i$ is the $i^{\text{th}}$ child of $C$. By definition of a DR-plan, it is implied that $C_i$ is a well-constrained vertex-maximal proper subgraph.

>>>>

Other generic constraint system or (hyper)graph properties are the
following:
% (\note we refer to these as properties of the constraint
% system or as properties of the underlying (hyper)graph
% interchangeably).
A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate
rigidity matrix of indeterminates has independent rows (i.e, the
determinant of some square submatrix is not identically zero).
%It is {\em rigid} if the number of independent rows of the rigidity matrix is
%at least the number of columns less the number of rigid body motions,
%which is 3 for distance constraint systems.
It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is
both rigid and independent.
%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$,
%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it
is independent and not rigid, or \dfn{overconstrained} if it is not
independent.

@@

\subsection{Decomposition-Recombination (DR) Plans}
%
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{overconstrained} if it is not independent.
%
%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
%
%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
%
% \begin{definition}
%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
%
%
%
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
    following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced
        subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
        \item A leaf node is a single vertex.
A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$
        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
        and in fact, $C$ contains all the edges in the union of its
        children as well.
        We denote a complete DR-plan of $G$ as $\comdrp{G}$
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
    The maximum  fan-in is called the \dfn{size} of the DR-plan.
    We denote an optimal DR-plan of $G$ as $\optdrp{G}$.

@@

% \begin{definition}
%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$. By this, rule 3 is implicit for a complete DR-plan.
% \end{definition}

@@

%\emph{wellconstrained} bar and joint graph with values $(V,E,w_V,w_E)$.
% \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages. It is not useful to solve for two bodies simultaneously if there are no constraints between them.
% THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.

@@

\medskip\noindent
\header{Realizing given qusecs.}
Note that this basic task
would require finding real solutions a large multivariate polynomial
system (of inequalities and equalities representing the constraints).
Without DR-planning, this requires double exponential time in the
number of variables (even if orientation type is specified). With
DR-planning, the complexity is dominated by the size of the largest
subsystem that is solved, or recombined from the solutions of its
child subsystems i.e.,the maximum fan-in occurring in a DR-plan. Thus
finding an optimal DR-plan is important for realization.

\medskip\noindent
\header{Recursive block
decomposition of the rigidity matrix, stress space and external stresses.}
This can be obtained using the DR-plan we obtain a
with each block being the
rigidity matrix of an isostatic subgraph. The blocks intersect on
isostatic or trivial subgraphs. This recursive block decomposition of
the rigidity matrix $R$ correspondingly recursively decomposes
the stress vectors $s$ that balance a given external stress $t$ since
$sR = t$. Moreover, the external stress vector $t$ is
recursively ``distributed'' into external stresses for each
node of the DR-plan.
When the rigidity matrix and stress
vectors contain indeterminates, this  can be used to design
underlying graph and parameters $\delta$ to obtain an optimal
distribution of stresses on the geometric primitives. When the
realization of qusecs is given, this can be used to analyze the
distribution of a given external load.

@@

%%    \item \textbf{overconstrained} if it is not independent.

@@

%%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.

@@

%%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.

@@

%%    \item \textbf{underconstrained} if $G$ is independent and not rigid.

@@

%%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.

@@

%\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must overconstrained and (2) all of its subgraphs are also trivial.

@@

%%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some

@@

%%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all

@@

%%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.

@@

%        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.

@@

%    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.

@@

%    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$. This makes rule 3 implicit. The complete DR-plan is unique.

@@

%%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$. By this, rule 3 is implicit for a complete DR-plan.

@@

%In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{wellconstrained} bar and joint graph with values $(V,E,w_V,w_E)$.

@@

%$C$ is the graph in an arbitrary node in $CompleteDRP(G)$. $C_i$ is the $i^{\text{th}}$ child of $C$. By definition of a DR-plan, it is implied that $C_i$ is a wellconstrained vertex-maximal proper subgraph.
